---
title: |
  Quality check of the personal data in 'Edad con Salud', Cohort 2011 - wave 4
output:
  word_document:
    df_print: kable
params:
  password: 'password for the contact data file'
---

```{r setup}
#| include: false

knitr::opts_chunk$set(
  results = 'asis',
  echo    = FALSE
)
```

```{r libraries}
#| include: false
library(tidyverse)
library(xlsx)
library(readxl)
library(ecs.data)
library(haven)
library(assertive.strings)
library(assertive.base)
library(glue)
library(lubridate)
```

# Introduction

This document shows the result of the quality checks performed on the personal
data (i.e. contact information) of the participants in the 4th wave of the 2011
cohort (i.e. *original cohort*) of the *Edad con Salud* project.

```{r load-data}
#| include: false

# The contact data is in a local path (and password-protected).
MAIN_DATASET_PATH   <- r"(~\Workspace\21057906DatosContacto20230116.xlsx)"
RETMAIL_DATA_PATH   <- file.path(
  read_ecs_folder("DOC"),
  "Edad con salud - Ola 4",
  "Participantes",
  "Cohorte 2011",
  "Incidencias muestra de seguimiento.xlsx"
)
FIELDWORK_DATA_PATH <- file.path(
  read_ecs_folder("DOC"),
  "Edad con salud - Ola 4",
  "Bases de datos",
  "Cohorte 2011",
  "Progreso del trabajo de campo",
  "Final",
  "20230116-FieldProgressReport-CálculoRatio.xlsx"
)

contact_data <- read.xlsx(
  MAIN_DATASET_PATH,
  sheetIndex = 1,
  password   = if (interactive()) rstudioapi::askForPassword()
               else               params$password
) |>
  as_tibble()

returned_mail_motives <- read.xlsx(
  RETMAIL_DATA_PATH,
  sheetIndex = "Cartas devueltas"
)             |>
  as_tibble() |>
  filter(year(FECHA.DEVOLUCIÓN) == 2022) # Keep only letters in last delivery

fieldwork_data <- read_excel(
  FIELDWORK_DATA_PATH,
  sheet = "Hoja1",
  skip  = 8
) |>
  rename(`Fieldwork result code` = `Final Result Code Household (Q0007a)...8`)

result_codes <- read_excel(
  FIELDWORK_DATA_PATH,
  sheet = "CÓDIGOS COURAGE",
  range = 'F6:F25'
) |>
  separate(
    `Final Result Code`,
    into    = c("Fieldwork result code", "Fieldwork result"),
    sep     = '= ?',
    convert = TRUE
  )

fieldwork_data <- fieldwork_data |>
  left_join(result_codes, by = "Fieldwork result code")
```

```{r variable-definitions}
#| include: false

VARIABLES <- contact_data                  |>
  colnames()                               |>
  enframe(name = NULL, value = "variable") |>
  add_column(
    name = c(
      "ID",
      "alternative ID",
      "proxy",
      "name",
      "surname",
      "proxy's name",
      "proxy's surname",
      "address",
      "town",
      "zip code",
      "comments",
      "ID nº",
      "Q0301a",
      "Q0303",
      "phone nº",
      "alternative phone nº",
      "Q0307",
      "Q0306",
      "Q0306b",
      "Q0306c", 
      "Q0306d",
      "NA."
    ),
    .before = 1
  ) |>
  deframe()

ID_VARS   <- VARIABLES[c("ID", "alternative ID")] |> syms() |> unname()
SURNAME   <- VARIABLES["surname"]
NAME      <- VARIABLES["name"]
ADDRESS   <- VARIABLES["address"]
PHONE     <- VARIABLES["phone nº"]
PHONE_ALT <- VARIABLES["alternative phone nº"]
ZIP_CODE  <- VARIABLES["zip code"]
TOWN      <- VARIABLES["town"]

# These codes are:
#   07=Unable to locate household or individual respondent
#   11=Unsafe or dangerous area or no access to household/indivudual respondent
#   21= Address doesn't exist
ADDRESS_RELATED_CODES <- c(7, 11, 21) # TODO: Consider code 10?
```

# Quality checks

The following data quality checks are performed on the contact information:

## Participants' names

### Completeness

Checks whether the surnames (variable ``r SURNAME``) are filled in.

```{r surnames}
surname_sym <- sym(SURNAME)

contact_data                                           |>
  filter(is_missing_or_empty_character(!!surname_sym)) |>
  select(!!!ID_VARS, !!surname_sym)
```

There are no empty values in ``r SURNAME``.

### Duplicates

Checks that the full names (variables ``r NAME`` and ``r SURNAME`` concatenated)
are not duplicated.

```{r name-duplicates}
name_sym <- sym(NAME)

contact_data               |>
  unite(
    "full_name",
    !!name_sym, !!surname_sym,
    sep    = ' ',
    remove = FALSE)        |>
  group_by(full_name)      |>
  mutate(occurences = n()) |>
  ungroup()                |>
  filter(occurences > 1)   |>
  select(!!!ID_VARS, !!name_sym, !!surname_sym)
```

There are no duplicates in the full names.

## Addresses

### Duplicates

Checks whether the addresses (variable ``r ADDRESS``) are not duplicated.

```{r address-duplicates}
address_sym <- sym(ADDRESS)

contact_data               |>
  group_by(!!address_sym)  |>
  mutate(occurences = n()) |>
  ungroup()                |>
  filter(occurences > 1)   |>
  select(!!!ID_VARS, !!address_sym)
```

There are no duplicates in ``r ADDRESS``.

### Returned letters

Checks the address of the cases that correspond to returned letters with
"incorrect address" or "incomplete address". The address in the letter and the
contact information must be the same. For the letters with an address-related
incidence, the addresses in the letters are manually checked.

```{r returned-letters}
town_sym <- sym(TOWN)
zipcode_sym <- sym(ZIP_CODE)

DIR_PATTERN <- "(?i)fallecid[oa]"

returned_mail_motives <- returned_mail_motives |>
  filter(ASUNTO |> str_detect(DIR_PATTERN, negate = TRUE))

wrong_dir_data <- contact_data                                        |>
  inner_join(returned_mail_motives, by = c(Q0002 = "ID"))             |>
  select(!!!ID_VARS, !!address_sym, !!zipcode_sym, !!town_sym, ASUNTO) |>
  arrange(Q0002)

LETTERS_MISMATCHING_DIR <- tribble(
  ~Q0002, ~mismatch,
     281,      TRUE,
     549,      TRUE,
    1772,     FALSE,
    2262,      TRUE,
    2331,     FALSE,
    2480,     FALSE,
    2541,     FALSE,
    2798,     FALSE,
    2856,     FALSE,
    2871,     FALSE,
    2941,     FALSE,
    2943,      TRUE,
    2948,     FALSE,
    2971,     FALSE,
    3024,     FALSE,
    3357,     FALSE,
    3449,     FALSE,
    3896,     FALSE, 
    4670,     FALSE,
    4675,     FALSE
)

wrong_dir_data <- wrong_dir_data |>
  left_join(LETTERS_MISMATCHING_DIR, by = "Q0002")
```

The following addresses do not match the addresses in the letters returned;
**please double check whether the participant has moved out to the new address
listed below**:

```{r returned-letters-mismatches}
wrong_dir_data     |>
  filter(mismatch) |>
  select(-mismatch, -ASUNTO)
```

The following addresses match the addresses in the letters returned;
**please double check whether the address listed below is correct, or update it
otherwise**:

```{r returned-letters-matches}
wrong_dir_data      |>
  filter(!mismatch) |>
  select(-mismatch)
```

### Complete address

Checks that the variable ``r ADDRESS`` is complete.
In order to this, we first separate the street (type), the number, and the rest
of the information.

```{r complete-address}
# Patterns to extract the address components:
STREETNUM_PATTERN  <- "(\\d{1,3}|S\\/N|SD/N)"
SEPARATOR_PATTERN  <- "[\\s,\\.ºª\\-N]"
STREETNAME_PATTERN <- glue("[^\\d]*(?={SEPARATOR_PATTERN}{STREETNUM_PATTERN})")
EXTRA_PATTERN      <- glue("(?<={STREETNUM_PATTERN})[^\\d].+")

split_addresses <- contact_data |> transmute(
  !!!ID_VARS, !!address_sym,
  address_num   = (!!address_sym) |> str_extract(STREETNUM_PATTERN),
  address_name  = (!!address_sym) |> str_extract(STREETNAME_PATTERN),
  address_extra = (!!address_sym) |> str_extract(EXTRA_PATTERN)
)
```

The following addresses are incomplete, because **they do not have an address
number, and thus need to be double checked**:

```{r incomplete-address-out}
split_addresses                                   |>
  filter(if_any(address_num:address_name, is.na)) |>
  select(!!!ID_VARS, !!address_sym)
```

<!-- The following addresses only have street name and number, so they may be
incomplete and need to be double checked: -->

```{r incomplete-address-out-extra}
#| include: false
#| echo:    true
#| results: hide
split_addresses                    |>
  filter(address_extra |> is.na()) |>
  select(!!!ID_VARS, !!address_sym)
## TODO: Filter out the ones that are known to be single houses
```

### Empty zip codes

Checks that there are no empty zip codes (variable ``r ZIP_CODE``).

```{r zip-codes}
contact_data                                           |>
  filter(is_missing_or_empty_character(!!zipcode_sym)) |>
  select(!!!ID_VARS, !!zipcode_sym)
```

There are no empty values in ``r ZIP_CODE``.

### Empty towns

Checks that there are no empty towns (variable ``r TOWN``).

```{r towns}
contact_data                                        |>
  filter(is_missing_or_empty_character(!!town_sym)) |>
  select(!!!ID_VARS, !!town_sym)
```

There are no empty values in ``r TOWN``.

## Telephone numbers

### Duplicates

Checks that the telephone numbers (variables ``r PHONE`` and ``r PHONE_ALT``)
are not duplicated.

```{r phone-duplicates}
phone_sym    <- sym(PHONE)
phonealt_sym <- sym(PHONE_ALT)

id_var <- ID_VARS[[1]]

contact_data                                               |>
  select(!!!ID_VARS, !!phone_sym, !!phonealt_sym)          |>
  pivot_longer(cols = c(!!phone_sym, !!phonealt_sym))      |>
  filter(value |> is_not_na())                             |>
  group_by(value)                                          |>
  mutate(occurences = n())                                 |>
  filter(occurences > 1, any(!!id_var != first(!!id_var))) |>
  pivot_wider(id_cols = c(!!!ID_VARS))
```

**The two cases shown have the same telephone number and need to be double
checked**.
