---
title: "Edad con Salud C2011-W4, check of (possible) incomplete addresses"
output: html_notebook
params:
  password: ''
---

# Introduction

This notebook helps perform a manual check of the addresses that may be
incomplete in the contact information of the participants in the 4th wave of the
2011 cohort of the Edad con Salud project.

The dataset file needs to be copied into a local folder
**outside the Rstudio project**.


**DO THE FOLLOWING STEPS:**

1. Make sure the dataset is in your file system, then
**update the path to the folder where the dataset file is located**, assigned to
object `DATASETS_FOLDER` in line 36.

2. Check that **the dataset file name coincides with the string in line 37**,
assigned to the object `MAIN_DATASET_PATH`. If it does not,
**update it with the proper name**.

3. Run chunk `setup`.

```{r setup}
library(tidyverse)
library(xlsx)
library(rstudioapi)
library(glue)

DATASETS_FOLDER   <- "~/Workspace"
MAIN_DATASET_FILE <- "21057906DatosContacto20230116.xlsx"
OUTPUT_FILE       <- "incomplete_addresses.csv"

MAIN_DATASET_PATH <- file.path(DATASETS_FOLDER, MAIN_DATASET_FILE)
OUTPUT_PATH       <- file.path(DATASETS_FOLDER, OUTPUT_FILE)
```

# Procedure

## Load personal data

Personal data must be first loaded.
When running the chunk `load-data`, you will be prompted to introduce the
dataset file password.

**DO THE FOLLOWING STEPS:**

1. Run chunk `load-data`.

2. **Input the access password** in the text box.

```{r load-data}
contact_data <- read.xlsx(
  MAIN_DATASET_PATH,
  sheetIndex = 1,
  password   = if (interactive()) askForPassword()
               else               params$password
) |>
  as_tibble()

VARIABLES <- contact_data                  |>
  colnames()                               |>
  enframe(name = NULL, value = "variable") |>
  add_column(
    name = c(
      "ID",
      "alternative ID",
      "proxy",
      "name",
      "surname",
      "proxy's name",
      "proxy's surname",
      "address",
      "town",
      "zip code",
      "comments",
      "ID nº",
      "Q0301a",
      "Q0303",
      "phone nº",
      "alternative phone nº",
      "Q0307",
      "Q0306",
      "Q0306b",
      "Q0306c",
      "Q0306d",
      "NA."
    ),
    .before = 1
  ) |>
  deframe()

ID_VARS   <- VARIABLES[c("ID", "alternative ID")] |> syms() |> unname()
ADDRESS   <- VARIABLES["address"]
ZIP_CODE  <- VARIABLES["zip code"]
TOWN      <- VARIABLES["town"]

address_sym <- sym(ADDRESS)
town_sym    <- sym(TOWN)
zipcode_sym <- sym(ZIP_CODE)

STREETNUM_PATTERN  <- "(\\d{1,3}|S\\/N|SD/N)"
SEPARATOR_PATTERN  <- "[\\s,\\.ºª\\-N]"
STREETNAME_PATTERN <- glue("[^\\d]*(?={SEPARATOR_PATTERN}{STREETNUM_PATTERN})")
EXTRA_PATTERN      <- glue("(?<={STREETNUM_PATTERN})[^\\d].+")

split_addresses <- contact_data |> mutate(
  !!!ID_VARS, !!address_sym,
  address_num   = (!!address_sym) |> str_extract(STREETNUM_PATTERN),
  address_name  = (!!address_sym) |> str_extract(STREETNAME_PATTERN),
  address_extra = (!!address_sym) |> str_extract(EXTRA_PATTERN),
  collapsed_address = paste(!!address_sym, !!town_sym, !!zipcode_sym)
)

incomplete_addresses <- split_addresses |> filter(address_extra |> is.na())

if (!file.exists(OUTPUT_PATH)) {

  incomplete_addresses                                          |>
    mutate(index = row_number(), incomplete = NA, remarks = NA) |>
    select(
      index, !!!ID_VARS,
      !!address_sym, !!town_sym, !!zipcode_sym,
      incomplete, remarks
    )                                                           |>
    write_excel_csv2(OUTPUT_PATH, na = '')
}
```

Chunk `load-data` will load the dataset, then it will perform some operations
to create the variables needed to do the search.
This includes creating a CSV file that will be filled in with the results of the
search (IF it doesn't exist yet).

## Create search index and iteration function

The following chunk creates a function that iterates through the cases,
outputs address of the the current case to the document output pane,
**pastes it to the clipboard**, and iterates the index to the next case.

**DO THE FOLLOWING STEPS:**

1. Open the file ``r OUTPUT_PATH``.

2. If it is the first time the procedure is run, go to step 2. If it has been
run before and some of the addresses have already been searched,
**update the numeric value of object `index` in line 160 to the first case**
**that has not been searched yet** (look for value _"CONTINUE HERE"_ in column
`remarks`).

3. Run chunk `create-iteration-function`.

```{r create-iteration-function}
index <- 1

next_address <- function() {

  URL <- "https://www.google.es/maps/place/"

  address <- incomplete_addresses |> slice(index) |> pull()

  paste0(URL, address) |> writeClipboard()
  cat(address, sep = '\n')
  
  index <<- index + 1
}
```

## Do the address search

When the previous steps are completed, the search has been setup.
Now, you will need to perform the search of the addresses and fill in the file
``r OUTPUT_PATH`` with the incomplete addresses.
In order to this, you will iterate through the dataset,
**searching for the addresses in Google Maps** to check whether they are
incomplete or not.

- **Incomplete addressess** are flats or buildings with more than one address,
where a floor and/or door nº are missing.

- **Complete addresses** are usually unifamiliar houses, where a floor and a
door nº does not exist.

**DO THE FOLLOWING STEPS:**

1. Open the file ``r OUTPUT_PATH`` in MS Excel. Double click on the right side
of the column headers if necessary to show the whole column span.

2. Open a web browser (Chrome, Safari, Firefox, etc.).

3. Run chunk `next-address` by going to line 244 and typing
`CTRL + SHIFT + ENTER` (Windows) or `CMD + SHIFT + ENTER` (Mac).

4. Type `ALT + TAB` (Windows) or `CMD + TAB` (Mac) until you change to the web
browser window.

5. Type `CTRL + L` (Firefox, Chrome, Edge) or `???` (Safari) to go to the URL
text box.

6. Type `CTRL + V` (Windows) or `CMD + V` (Mac) to paste in the search URL.

7. If this automatic search fails, please introduce manually the address in the
search textbox in Google Maps.

8. Check the Street View preview to
**find out whether the address corresponds to a complete or an incomplete one**.

9. If you determine the address is:

   - **Incomplete**: Go to the corresponding row in ``r OUTPUT_PATH`` and
   type in "TRUE" in column G ("incomplete").
   
   - **You can't decide**: Go to the corresponding row in ``r OUTPUT_PATH`` and
   explain why in column H ("remarks").
   
   - **Complete**: Do nothing.
   
10. Go to step 3, until you decide to stop the session or have finished with all
    the cases.

### IMPORTANT ADDITIONAL STEP

If you decide to stop the session

**DO THE FOLLOWING STEPS:**

1. Go to the row corresponding to **the next case to check** when you restart
the session.

2. Type in _"CONTINUE HERE"_ in column H ("remarks").

3. Save the file in MS Excel; when prompted whether
"to keep using the CSV format" choose "YES".

4. Close Excel.

```{r next-address}
next_address()
```

